Exploratory Analysis

### 2025-12-04 Exploratory Analysis.Rmd
Basic intro plots & finding outliers
Tried clustering

>>>>

```{r}

df1 = read.csv("RWAR_temp_pH_cond_turb_cleaned.csv")
df2 = read.csv("alkalinity_hardness_cleaned.csv")
df3 = read.csv("CoagulantDoses_2017-20.csv")

df1 = df1 %>%
  mutate(Date = ymd(Timestamp))
df2 = df2 %>%
  mutate(Date = mdy(Date))
df3 = df3 %>%
  mutate(Date = mdy(Date))

# This merge is only keeping the rows that are in both data frames
# Not using df3 because I don't want our target variables to be included in the clustering
merged = df1 %>% inner_join(df2, by = "Date")
merged$Date = NULL # We already have the Timestamp column

cluster_data = merged %>%
  select(where(is.numeric)) %>%
  drop_na()

cluster_data = scale(cluster_data) # Standardize the data

cluster_fit = kmeans(cluster_data, centers = 5, nstart = 25)

merged$cluster = cluster_fit$cluster

# Interpretation

table(merged$cluster)
cluster_fit$centers

ggplot(merged, aes(x = cluster_data[ , "mean_Temp"], y = cluster_data[ , "mean_pH"], # you can compare any two columns
                   color = factor(cluster))) +
  geom_point() +
  labs(title = "K-Means Clustering", color = "Cluster") +
  theme_bw()
